Declarative runtime reconfiguration of pulverised distributed systems \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-11 paper-2023-acsos-pulverization.tex
better introduce pulverisation and glue with the DSL part

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{acronym}
\usepackage{listing-style}
\usepackage[hidelinks]{hyperref}

\acrodef{dsl}[DSL]{domain-specific language}

@@

Previous work suggests a technique for decoupling business logic and deployment concerns
called \emph{pulverization}~\cite{FI2020-pulverization}.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of five micro-components
(behaviour, state, communication, sensors, and actuators)
whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changing infrastructures or non-functional requirements.
%
One limitation of the current pulverisation approach regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components in face of changing conditions or requirements.

@@

\subsection{Domain model}

\subsection{Domain-specific language}

The framework allows the system configuration via two distinct \acp{dsl}, one for defining the system's devices and their components,
and one for defining the runtime configuration of the system.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{acronym}
\usepackage{listing-style}
\usepackage[hidelinks]{hyperref}
\usepackage[inline]{enumitem}

\acrodef{dsl}[DSL]{domain-specific language}
\acrodef{iot}[IoT]{Internet of Things}

@@

Previous work suggests a technique for decoupling business logic and deployment concerns
called \emph{pulverization}~\cite{FI2020-pulverization}.
%
In short, the core idea is to consider devices as \emph{logical} entities,
whose software can be designed as (or broken down to, if already existing) an ensemble of five micro-components
(behaviour, state, communication, sensors, and actuators)
whose actual host can be decided at deployment time.
%
This way, the designer can focus on the business logic at hand,
delaying an optimised deployment to later stages of the sofware design,
thus gaining resilience to changes in the infrastructure or in non-functional requirements.
%
One limitation of the current pulverisation approach regards dynamicity:
although the application can be deployed on arbitrary systems,
there is no support for the reconfiguration of components in face of changing conditions or requirements.

@@

The goal of pulverisation~\cite{FI2020-pulverization}
is to provide application designers a way to specify the functional semantics of their software in a deployment-independent way.
%
To do so,
the application logic should be designed considering a \emph{logical system},
namely, a network of \emph{logical} devices forming a network with \emph{arbitrary} topology.
%
For a pulverisation system to work,
the application of every single logical device must be decomposable into an ensemble of \emph{pulverised components}
representing, respectively:
\begin{itemize}
    \item a set $\sigma$ of logical sensors;
    \item a set $\alpha$ of logical actuators;
    \item a state $\kappa$, representing the logical device's knowledge;
    \item a communication component $\chi$,
    handling interaction with reachable devices in the logical system, and
    \item a computation component $\beta$, modelling the behaviour of the logic device.
\end{itemize}
%
Decomposition of an application into pulverised components can be achieved in two ways:
either the application is designed with pulverisation in mind,
or the application is developed using a framework supporting automatic decomposition
(one notable example are aggregate computing frameworks~\cite{BealIEEEComputer2015}).
%
Once the application has been pulverised,
a mapping must be provided between the pulverised logical system and the hosts that will execute the pulverised components.
%
In this process,
a single logical component could be (and usually does) end up being executed on multiple hosts.

For instance, let us assume a minimal logical system counting two devices:
two rain gauges logging the water level for future reference,
and opening a valve when the water level crosses a threshold in both devices.
%
Once pulverised, the logical system would be split in ten deployable pulverised components
(that we indicate with their component symbol and an index, e.g., $\sigma_1$ is the logical sensor of the first device).
%
These pulverised components can be deployed in arbitrary hosts,
as far as they have the \emph{capabilities} required to host them
(for instance, the component hosting a rain gauge logical sensor must have an actual sensor,
while the component hosting the behaviour must be sufficiently powerful to execute the logic).
%
Thus, with \emph{no change to the application logic},
the system could be deployed on very different systems:
let us assume, for instance,
that our actual target infrastructure is not composed of two connected devices,
but it is an existing \ac{iot} system
where the rain gauges are connected to two LoRaWAN motes
(which are too weak and energy-critical to execute the logic),
the valves are controlled by a dedicated device,
the levels must be logged on a database hosted on a cloud server reachable via HTTPS,
and we have an internal edge server that we can use as we please.
%
In this case, we would deploy $\sigma_1$ and $\sigma_2$ on the motes,
$\alpha_1$ and $\alpha_2$ on the valve controller,
$\kappa_1$ and $\kappa_2$ on the cloud server,
and the remaining components on the edge server.

\meta{maybe import the image from the pulverisation paper?}

\subsection{Domain-specific languages for pulverisation}

From the previous section,
we can identify two main phases in the pulverisation process:
\begin{enumerate*}[label=\it{(\roman*)}]
  \item the definition of the available devices and their capabilities; and
  \item the allocation of the pulverised components to them.
\end{enumerate*}
%
We decided to capture these phases into two distinct \acp{dsl}.
